---
title: "BELLABEAT CASE STUDY July 2022"
author: "Elena Olivo"
date:   "July 23, 2022"
output: html_notebook
---

#### Recent trends in smart device usage suggest health data collection by
#### users is an important feature with high engagement.


#### Bellabeat would like to engage their app users in new ways, by looking at
#### data from these smart devices to create new opportunities within the app.


#### The aim here is data exploration to discover insights that can benefit 
#### Bellabeat, and find new ways to engage their app users.





#### Load packages:
#### * Tidyverse
#### * ggplot
#### * dplyr
#### * skimr, janitor, here
#### * reshape




# Examine health dataset to begin analysis


```{r}
head(health_data)
```
#### Cleaning data


```{r}
clean_names(health_data)
```
#### Organize and examine data Body Mass Index (BMI)
<https://www.niddk.nih.gov/health-information/health-statistics/overweight-obesity>


```{r}
health_data %>% arrange(-BMI)
```
#### BMI by healthy ratio (under 24.9)


```{r}
filter(health_data, BMI <=24.9) 
```
#### BMI Analysis unhealthy level (over 25.0)


```{r}
filter(health_data, BMI > 25.0)
```
#### Examine and observations


```{r}
health_data %>% summarize(mean_BMI = mean(BMI))
```
#### Data examination


```{r}
health_data %>% summarize(max_BMI = max(BMI))
```
#### Analysis: Mean BMI 25.1, and (one outlier of 47.54). Target users by BMI greater than 24.9 (unhealthy BMI)?


```{r}
health_data %>% arrange(-BMI)
```
## New dataset


```{r}
health_data2 <- health_data %>% arrange(-BMI)
```


#### Examine with a plot


```{r}
ggplot(data= health_data2) +
  geom_point(mapping=aes(x=Id, y=BMI), color ="blue") +
  labs(title = "BMI LEVELS")
```
### One OUTLIER of BMI = 47.54

### Density plot for another look at data


```{r}
ggplot(health_data2, aes(x=BMI)) +
  geom_density(color = "darkblue", fill = "lightblue") +
  labs(title = "BMI Levels")
```


#### Analysis: Mean BMI levels 25. Target this level and higher.


## Sleep pattern analysis


```{r}
head(sleep_data)
```
#### Look at Sleep data


```{r}
clean_names(sleep_data)
```
#### Analyze by unique Id


```{r}
sleep_data %>% group_by(Id, SleepDay) %>% drop_na() %>% summarize(max_bed=max(TotalTimeInBed),
max_sleep=max(TotalMinutesAsleep))
```
#### Analyze data by ID


```{r}
user_0366 %>% filter(Id ==1503960366) 
```
#### Filter by Id


```{r}
head(user_1503)
```
#### Analysis: Make note that more time in bed than sleeping = Meditation exercise?
<https://www.healthline.com/health/meditation-for-sleep>


### Select unique ID (1503) for analysis with a scatterplot


```{r}
ggplot(user_1503, mapping=aes(x=max_bed, y=max_sleep)) +
  geom_point() +
  labs(title = "Comparison Sleep vs. Bed Time")
```
#### Create new dataframes for better plot (specific observations from month)


```{r}
x <- c(346, 407, 442, 367, 712, 320, 377, 364, 384, 449, 323, 274, 393, 354)
y <- c(327, 384, 412, 340, 700, 304, 360, 325, 361, 430, 277, 245, 366, 341)
day <- c(12, 13, 15, 16, 17, 19, 20, 21, 23, 24, 25, 26, 28, 29)
```


#### Step 2 of creating new dataframes


```{r}
df1 <- data.frame(x, y, day)
df2 <- melt(df1, id.vars='day')
head(df2)
```
### New plot style to better show comparison
#### (x = Max in-bed time, y = time asleep). 
#### Small set of objects from dataset used for visualization.


```{r}
ggplot(df2, aes(x=day, y=value, fill=variable)) +
    geom_bar(stat='identity', position='dodge') +
    labs(title = "MEDITATION BEFORE SLEEP OPPORTUNITY?") 
```


#### Analysis: Suggest meditation exercise through app (_audible_?)



## Correlation between steps and calories by day?


```{r observations}
nrow(steps_by_day)
nrow(calories_by_day)
```


#### Steps data (daily)


```{r}
steps_by_day %>%
  summary()
```
#### Calories data - daily


```{r}
calories_by_day %>%
  summary()
```
#### Clean data


```{r}
clean_names(steps_by_day)
```


#### Clean Data


```{r}
clean_names(calories_by_day)
```


#### Combine datasets


```{r}
stepscal_1 <-merge(steps_by_day, calories_by_day, by = c("Id", "ActivityDay"))
```


#### Look at combination dataset


```{r}
head(stepscal_1)
```
#### Arrange dataset
```{r}
arrange(stepscal_1, desc(StepTotal))
```
#### Any trends via plot?


```{r}
ggplot(data=stepscal_1) +
   geom_point(mapping=aes(x=StepTotal, y=Calories), color="purple") +
   labs(title = "Relation between steps and calories")
```


#### Another plot with smooth feature
```{r}
ggplot(stepscal_1, aes(x= StepTotal, y= Calories)) + 
  geom_point() +        
  geom_smooth(method="gam", 
formula = y ~s(x))
```


#### Analysis: More steps, more calories burned!


## New Steps/Calories analysis: Look at data, target by unique Id, patterns?


```{r}
glimpse(hourly_steps_v1)
```


#### Review data


```{r}
glimpse(hourly_calories_v1)
```


#### Merge datasets
```{r}
combined_dataV2 <-inner_join(hourly_calories_v1, hourly_steps_v1, by = c("Id", "ActivityHour"))
```


#### Filter by Id


```{r}
combined_dataV2 %>% filter(Id == "4702921684")
```


#### View data for creation of alert system of low activity points in day

```{r}
user_1684 %>%
  select(ActivityHour,
         Calories,
         StepTotal) %>%
  summary()
```


#### Subset of rows


```{r}
user_1684v2 <- user_1684[c(1:24), c(2,3:4)]
```


#### Arrange subset

```{r}
arrange(user_1684v2)
```


#### Plot


```{r}
ggplot(user_1684v2) +
  geom_point(mapping = aes(x=ActivityHour, y=StepTotal)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Tailor app to individual by daily data usage")
```


#### During lower points of activity during day, perhaps an opportunity 
#### to engage user to be more active via alert?


## New dataset on steps
```{r}
head(hourly_intense_active2)
```


#### Summary of dataset
```{r}
colnames(hourly_intense_active2)
```


#### Look at sample of 50 observatons
```{r}
summary(hourly_intense_active2, 50)
```


#### Plot for analysis
```{r}
ggplot(data=hourly_intense_active2) +
geom_point(mapping =aes (x=AverageIntensity, y=TotalIntensity))
 
```


#### Subset for patterns or trends by unique user Id
```{r}
user_9391 <-hourly_intense_active2 %>% filter(Id == "8877689391") 
```


#### Get summary data
```{r}
min(user_9391$ActivityHour)
```


#### Get additional summary data
```{r}
max(user_9391$ActivityHour)
```


#### Can this data be used to explain higher activity on a particular day/hour?


#### Additional look at unique Id
```{r}
user_9391 %>% arrange(ActivityHour)
```


#### Subset by Id and sample of data
```{r}
head(user_9391, 24) %>% arrange(ActivityHour)
```


#### View new subset dataframe
```{r}
head(user_9391v2)
```


#### Plot this subset
```{r}
ggplot(data = user_9391v2) +
  geom_point(mapping = aes(x=ActivityHour, y = TotalIntensity)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Identify patterns over weeks")
```


#### Another style of plot


```{r}
ggplot(data = user_9391v2) +
  geom_bar(mapping = aes(x=ActivityHour, fill = TotalIntensity)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title="Pattern analysis by levels over time")
```


#### Follow up survey to identify days/hours when activity is low vs. high and why?


# SUMMARY
#### * Target users via app by Body Mass Index (BMI): Diet and nutrition programs by BMI level?
#### * Alert users via app when BMI is approaching unhealthy level (_audible_) 
#### * Create meditation exercise through app for bedtime (_before_ / _after_)
#### * Target individuals via app to increase activity levels during low periods of activity (_day_/_week_)
#### * Use surveys via app to question users on periods of low activity vs. high activity?

## DATASET REFERENCE 
<https://www.kaggle.com/datasets/arashnic/fitbit>

## SMART WATCH DATA RESEARCH/TRENDS
<https://www.delveinsight.com/blog/wearable-technology-trends-2022>

<https://www.crn.com/slide-shows/internet-of-things/ces-2022-10-wearable-technologies-for-health-safety-fitness#.YpaevEdbhEE.link>























